Another paper with lots of text and fancy pictures

### 2023-11-27 paper.tex
Minor fixes

Now we can define the three-point correlation function
$S_3: \mathbb{R}^2 \rightarrow [0, 1]$:
\begin{equation}
  S_3^n(x_1, x_2) = \langle I_n(x) I_n(x + x_1) I_n(x + x_2) \rangle
\end{equation}
where $\langle \dots \rangle$ means average over all points in $A$. It can be
understood as the probability that all three vertices of a triangle
$(0, x_1, x_2)$ will appear in the subset $A_n$ when the triangle is randomly
thrown in $A$. When $A$ represent isotropic medium, $x_1$ and $x_2$ can be
replaced with three numbers $r_1 = |x_1|$, $r_2 = |x_2|$ and $r_3 = |x_2 - x_1|$
to get a function of three scalar arguments.

@@

The second approach is to compute the whole correlation map in Fourier
domain. Suppose we have a function
$f: \mathbb{R} \rightarrow \mathbb{R}$. Triple correlation for this function is
a function
\begin{equation}
  g(t_1, t_2) = \int f(\tau) f(\tau + t_1) f(\tau + t_2) d \tau
\end{equation}
We can use a well-known formula to compute Fourier image of $g$:
\begin{equation}
  \hat{g}(z_1, z_2) = \hat{f}(z_1) \hat{f}(z_2) \overline{\hat{f}(z_1 + z_2)}
\end{equation}
This algorithm has a better computational complexity compared with the previous
algorithms if the whole correlation map is needed ($O(n \log n)$ vs
$O(n^2)$). The whole correlation map, however, requires a lot of memory and is
rarely needed. Imagine, that the input array $A$ is of dimensions
$500 \times 500$. Then, considering that single precision floating point numbers
are used, $4 \cdot 2 \cdot 500^4 / 1000^3 = 500$ GB are needed to store Fourier
image of the map. Moreover, Fourier image of the whole map is needed even if
only one element of the map in spatial domain is required. The aforementioned
limitations render this approach very impractical.

@@

To test $F_{sss}$ function we use a recently developed approach which computes
that function for sets defined by inequation $f(x) \le T$, $x \in \mathbb{R}^3$
with help of automatic differentiation \cite{postnicov20232}. We choose $f$
\begin{equation}
  f(x, y, z) = [(2x)^{\frac{3}{2}} + y^{\frac{3}{2}} + z^{\frac{3}{2}}]^{\frac{2}{3}}
\end{equation}
so inequation $f(x, y, z) \le T$ gives us an ellipsoid-like object
(\cref{fig:sss-ellipsoid}).

@@

\section{Discussion}
In this paper we have provided two approaches to compute three-point correlation
function:
\begin{enumerate}
\item Computation in frequency domain using Fourier transform. This approach has
  computational complexity $O(n \log n)$ and computes the whole correlation
  map. Unfortunately, it requires $O(n^2)$ additional memory and therefore can
  be used only with small inputs. Moreover, the whole Fourier image of the input
  is needed even if we need only a small part of the correlation map.
\item Computation in spatial domain. This approach is implemented by us in
  CorrelationFunctions.jl package. This approach calculates the full map with
  complexity $O(n^2)$ but uses only $O(n)$ additional memory for shifted
  versions of the input. For our purposes the whole map is not needed and
  therefore the amount of calculations can be significantly reduced.
\end{enumerate}
It's tempting to create an algorithm which unites advantages of both approaches
(fast speed of the first and low memory consumption of the second) which
computes some desired slices of correlation map. It is known that Fourier
transform is the only linear transform which allows computation of convolution
in $O(n)$. It is still an open question if there are linear transforms which
allow computation of convolution with complexity between $O(n)$ and $O(n^2)$
\cite{stone2008uniqueness,stone1998convolution}.

>>>>

Now we can define the three-point correlation function
$S_3: \mathbb{R}^{2n} \rightarrow [0, 1]$:
\begin{equation}
  S_3^n(x_1, x_2) = \langle I_n(x) I_n(x + x_1) I_n(x + x_2) \rangle
\end{equation}
where $\langle \dots \rangle$ means average over all points in $A$. It can be
understood as the probability that all three vertices of a triangle
$(0, x_1, x_2)$ will appear in the subset $A_n$ when the triangle is randomly
thrown in $A$. When $A$ represent isotropic medium, $x_1$ and $x_2$ can be
replaced with three numbers $r_1 = |x_1|$, $r_2 = |x_2|$ and $r_3 = |x_2 - x_1|$
to get a function of three scalar arguments.

@@

The second approach is to compute the whole correlation map in Fourier
domain. Suppose we have a function
$f: \mathbb{R} \rightarrow \mathbb{R}$. Triple correlation for this function is
a function
\begin{equation}
  g(t_1, t_2) = \int f(\tau) f(\tau + t_1) f(\tau + t_2) d \tau
\end{equation}
We can use a well-known formula to compute Fourier image of $g$:
\begin{equation}
  \hat{g}(z_1, z_2) = \hat{f}(z_1) \hat{f}(z_2) \overline{\hat{f}(z_1 + z_2)}
\end{equation}
This algorithm has a better computational complexity compared with the previous
algorithms if the whole correlation map is needed ($O(n^2 \log n)$ vs
$O(n^3)$ where $n$ is a number of elements in the input array). The whole
correlation map, however, requires a lot of memory and is rarely
needed. Imagine, that the input array $A$ is of dimensions $500 \times 500$.
Then, considering that single precision floating point numbers are used,
$4 \cdot 2 \cdot 500^4 / 1000^3 = 500$ GB are needed to store Fourier image of
the map. Moreover, Fourier image of the whole map is needed even if only one
element of the map in spatial domain is required. The aforementioned limitations
render this approach very impractical.

@@

To test $F_{sss}$ function we use a recently developed approach which computes
that function for sets defined by inequation $f(x) \le T$, $x \in \mathbb{R}^3$
with help of automatic differentiation \cite{postnicov20232}. We choose $f$
\begin{equation}
  f(x, y, z) = (|2x|^{\frac{3}{2}} + |y|^{\frac{3}{2}} + |z|^{\frac{3}{2}})^{\frac{2}{3}}
\end{equation}
so inequation $f(x, y, z) \le T$ gives us an ellipsoid-like object
(\cref{fig:sss-ellipsoid}).

@@

\section{Discussion}
In this paper we have provided two approaches to compute three-point correlation
function:
\begin{enumerate}
\item Computation in frequency domain using Fourier transform. This approach has
  computational complexity $O(n^2 \log n)$ and computes the whole correlation
  map. Unfortunately, it requires $O(n^2)$ additional memory and therefore can
  be used only with small inputs. Moreover, the whole Fourier image of the input
  is needed even if we need only a small part of the correlation map.
\item Computation in spatial domain. This approach is implemented by us in
  CorrelationFunctions.jl package. This approach calculates the full map with
  complexity $O(n^3)$ but uses only $O(n)$ additional memory for shifted
  versions of the input. For our purposes the whole map is not needed and
  therefore the amount of calculations can be significantly reduced.
\end{enumerate}
It's tempting to create an algorithm which unites advantages of both approaches
(fast speed of the first and low memory consumption of the second) which
computes some desired slices of correlation map. It is known that Fourier
transform is the only linear transform which allows computation of convolution
in linear time. It is still an open question if there are linear transforms which
allow computation of convolution faster than a naïve approach
\cite{stone2008uniqueness,stone1998convolution}.
